% Variable Neighborhood Search

% The Clever Algorithms Project: http://www.CleverAlgorithms.com
% (c) Copyright 2010 Jason Brownlee. Some Rights Reserved. 
% This work is licensed under a Creative Commons Attribution-Noncommercial-Share Alike 2.5 Australia License.

\documentclass[a4paper, 11pt]{article}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{url}
\usepackage[pdftex,breaklinks=true,colorlinks=true,urlcolor=blue,linkcolor=blue,citecolor=blue,]{hyperref}
\usepackage{geometry}
\usepackage[ruled, linesnumbered]{../algorithm2e}
\usepackage{listings} 
\usepackage{textcomp}
\ifx\pdfoutput\@undefined\usepackage[usenames,dvips]{color}
\else\usepackage[usenames,dvipsnames]{color}
\lstset{basicstyle=\footnotesize\ttfamily,numbers=left,numberstyle=\tiny,frame=single,columns=flexible,upquote=true,showstringspaces=false,tabsize=2,captionpos=b,breaklines=true,breakatwhitespace=true,keywordstyle=\color{blue},stringstyle=\color{ForestGreen}}
\geometry{verbose,a4paper,tmargin=25mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}

% Dear template user: fill these in
\newcommand{\myreporttitle}{Variable Neighborhood Search}
\newcommand{\myreportauthor}{Jason Brownlee}
\newcommand{\myreportemail}{jasonb@CleverAlgorithms.com}
\newcommand{\myreportwebsite}{http://www.CleverAlgorithms.com}
\newcommand{\myreportproject}{The Clever Algorithms Project\\\url{\myreportwebsite}}
\newcommand{\myreportdate}{20100208}
\newcommand{\myreportversion}{1}
\newcommand{\myreportlicense}{\copyright\ Copyright 2010 Jason Brownlee. Some Rights Reserved. This work is licensed under a Creative Commons Attribution-Noncommercial-Share Alike 2.5 Australia License.}

% leave this alone, it's templated baby!
\title{{\myreporttitle}\footnote{\myreportlicense}}
\author{\myreportauthor\\{\myreportemail}\\\small\myreportproject}
\date{\today\\{\small{Technical Report: CA-TR-{\myreportdate}-\myreportversion}}}
\begin{document}
\maketitle

% write a summary sentence for each major section
\section*{Abstract} 
% project
The Clever Algorithms project aims to describe a large number of Artificial Intelligence algorithms in a complete, consistent, and centralized manner, to improve their general accessibility. 
% template
The project makes use of a standardized algorithm description template that uses well-defined topics that motivate the collection of specific and useful information about each algorithm described.
% report
This report describes the Variable Neighborhood Search algorithm using the standardized template.

\begin{description}
	\item[Keywords:] {\small\texttt{Clever, Algorithms, Description, Optimization, Variable, Neighborhood, Search}}
\end{description} 

\section{Introduction} 
\label{sec:intro}
% project
The Clever Algorithms project aims to describe a large number of algorithms from the fields of Computational Intelligence, Biologically Inspired Computation, and Metaheuristics in a complete, consistent and centralized manner \cite{Brownlee2010}.
% description
The project requires all algorithms to be described using a standardized template that includes a fixed number of sections, each of which is motivated by the presentation of specific information about the technique \cite{Brownlee2010a}.
% this report
This report describes the Variable Neighborhood Search algorithm using the standardized template.

% Name
% The algorithm name defines the canonical name used to refer to the technique, in addition to common aliases, abbreviations, and acronyms. The name is used in terms of the heading and sub-headings of an algorithm description.
\section{Name} 
\label{sec:name}
% What is the canonical name and common aliases for a technique?
% What are the common abbreviations and acronyms for a technique?
% The heading and alternate headings for the algorithm description.
Variable Neighborhood Search, VNS

% Taxonomy: Lineage and locality
% The algorithm taxonomy defines where a techniques fits into the field, both the specific subfields of Computational Intelligence and Biologically Inspired Computation as well as the broader field of Artificial Intelligence. The taxonomy also provides a context for determining the relation- ships between algorithms. The taxonomy may be described in terms of a series of relationship statements or pictorially as a venn diagram or a graph with hierarchical structure.
\section{Taxonomy}
\label{sec:taxonomy}
% what is it
its a metaheuristic, it a global search technique that embeds a local search technique
% related
related to iterative local search

% Strategy: Problem solving plan
% The strategy is an abstract description of the computational model. The strategy describes the information processing actions a technique shall take in order to achieve an objective. The strategy provides a logical separation between a computational realization (procedure) and a analogous system (metaphor). A given problem solving strategy may be realized as one of a number specific algorithms or problem solving systems. The strategy description is textual using information processing and algorithmic terminology.
\section{Strategy}
\label{sec:strategy}
% What is the information processing objective of a technique?
% What is a techniques plan of action?

iterative explores neighborhoods of growing size - locates a local optima, keeps searching and expanding neighborhood size until an improved local optima is found, then the process is repeated

strategy is all about search sub-spaces with both stochastic and deterministic elements

three principles: (quoted from any review paper) A local minimum with respect to one neighbourhood structure is not necessary so for another; A global minimum is a local minimum with respect to all possible neighbourhood structures.  For many problems local minima with respect to one or several neighbourhoods are relatively close to each other.

the benefit of VNS seems to come from the fact that local optima are empirically commonly close to global optima n some problems under some neighbourhood types 

% Procedure: Abstract computation
% The algorithmic procedure summarizes the specifics of realizing a strategy as a systemized and parameterized computation. It outlines how the algorithm is organized in terms of the data structures and representations. The procedure may be described in terms of software engineering and computer science artifacts such as pseudo code, design diagrams, and relevant mathematical equations.
\section{Procedure}
\label{sec:procedure}
% What is the computational recipe for a technique?
% What are the data structures and representations used in a technique?
Algorithm~\ref{alg:variable_neighborhood_search} provides a pseudo-code listing of the Variable Neighborhood Search algorithm for minimizing a cost function.
% explaination
The pseudo code shows that the series of neighborhoods are search for each change to the local optimum. A systematic search of expanding neighborhoods for a local optimum is abandoned when an global improvement is achieved, shown in the pseudo code with the \texttt{Break} jump.

\begin{algorithm}[ht]
	\SetLine
	% data
	\SetKwData{Neighborhoods}{Neighborhoods}
	\SetKwData{CurrentNeighborhood}{$Neighborhood_{i}$}
	\SetKwData{Neighborhood}{$Neighborhood_{curr}$}
	\SetKwData{Best}{$S_{best}$}
	\SetKwData{Candidate}{$S_{candidate}$}
	% functions
	\SetKwFunction{Break}{Break}
	\SetKwFunction{Cost}{Cost}
	\SetKwFunction{StopCondition}{StopCondition}
	\SetKwFunction{CalculateNeighborhood}{CalculateNeighborhood}
	\SetKwFunction{RandomSolution}{RandomSolution}
	\SetKwFunction{RandomSolutionInNeighborhood}{RandomSolutionInNeighborhood}
  	\SetKwFunction{LocalSearch}{LocalSearch}

	% I/O
	\KwIn{\Neighborhoods}
	\KwOut{\Best}
  	% Algorithm
	\Best $\leftarrow$ \RandomSolution{}\;
	\While{$\neg$ \StopCondition{}} {
		\ForEach{\CurrentNeighborhood $\in$ \Neighborhoods} {
			% determine neighbourhood
			\Neighborhood $\leftarrow$ \CalculateNeighborhood{\Best, \CurrentNeighborhood}\;
			% starting point
			\Candidate $\leftarrow$ \RandomSolutionInNeighborhood{\Neighborhood}\;
			% local search
			\Candidate $\leftarrow$ \LocalSearch{\Candidate}\;
			% keep best
			\If{\Cost{\Candidate} $<$ \Cost{\Best}} {
				\Best $\leftarrow$ \Candidate\;
				\Break\;
			}
		}
	}
	\Return{\Best}\;
	% caption
	\caption{Pseudo Code Listing for the Variable Neighborhood Search algorithm.}
	\label{alg:variable_neighborhood_search}
\end{algorithm}

% Heuristics: Usage guidelines
% The heuristics element describe the commonsense, best practice, and demonstrated rules for applying and configuring a parameterized algorithm. The heuristics relate to the technical details of the techniques procedure and data structures for general classes of application (neither specific implementations not specific problem instances). The heuristics are described textually, such as a series of guidelines in a bullet-point structure.
\section{Heuristics}
\label{sec:heuristics}
% What are the suggested configurations for a technique?
% What are the guidelines for the application of a technique to a problem instance?
\begin{itemize}
	\item The Variable Neighborhood Descent (VND) can be embedded in the Variable Neighborhood Search as a the Local Search procedure and has been shown to be most effective. 
	\item Approximation methods (such as stochastic hill climbing) are suggested for use as the Local Search procedure for large problem instances in order to reduce the running time.
	\item Variable Neighborhood Search has been applied to a very wide array of combinatorial optimization problems as well as clustering and continuous function optimization problems.
	\item The embedded Local Search technique should be specialized to the problem type and instance to which the technique is being applied. 
\end{itemize}

% The code description provides a minimal but functional version of the technique implemented with a programming language. The code description must be able to be typed into an appropriate computer, compiled or interpreted as need be, and provide a working execution of the technique. The technique implementation also includes a minimal problem instance to which it is applied, and both the problem and algorithm implementations are complete enough to demonstrate the techniques procedure. The description is presented as a programming source code listing.
\section{Code Listing}
\label{sec:code}
% How is a technique implemented as an executable program?
% How is a technique applied to a concrete problem instance?
Listing~\ref{variable_neighborhood_search} provides an example of the Variable Neighborhood Search algorithm implemented in the Ruby Programming Language. 
% probem
todo

i'm thinking a TSP with 2-opt

% the listing
\lstinputlisting[firstline=7,language=ruby,caption=Variable Neighborhood Search algorithm in the Ruby Programming Language, label=variable_neighborhood_search]{../../src/algorithms/stochastic/variable_neighborhood_search.rb}


% References: Deeper understanding
% The references element description includes a listing of both primary sources of information about the technique as well as useful introductory sources for novices to gain a deeper understanding of the theory and application of the technique. The description consists of hand-selected reference material including books, peer reviewed conference papers, journal articles, and potentially websites. A bullet-pointed structure is suggested.
\section{References}
\label{sec:references}
% What are the primary sources for a technique?
% What are the suggested reference sources for learning more about a technique?

% 
% Primary Sources
% 
\subsection{Primary Sources}
% initial
Variable Neighborhood Search was initially proposed in an abstract by Mladenovic \cite{Mladenovic1995}.
% names
The approach is explained in terms of three different variations on the general theme. Variable Neighborhood Descent (VND) refers to the use of a Local Search procedure and the deterministic (as opposed to stochastic or probabilistic) change of neighborhood size. Reduced Variable Neighborhood Search (RVNS) involves performing a stochastic random search within a neighborhood and no refinement via a local search technique. Basic Variable Neighborhood Descent is the canonical approach described by Mladenovic and Hansen in the seminal paper on the approach \cite{Mladenovic1997}.

% 
% Learn More
% 
\subsection{Learn More}
% overview
Hansen and Mladenovic provide an overview of the approach that includes its recent history, extensions and a detailed review of the numerous areas of application \cite{Hansen2003}.

% extensions
There are many extensions to Variable Neighborhood Search, not limited to the following: Variable Neighborhood Decomposition Search (VNDS) involves embedding a second heuristic or metaheuristic approach in VNS to replace the Local Search procedure \cite{Hansen2001}, Skewed Variable Neighborhood Search (SVNS) encourages exploration of neighborhoods far away from discovered local optima, and Parallel Variable Neighborhood Search (PVNS) that either parallelize the local search of a neighborhood or parallelize the searching of the neighborhood themselves.


NOTES

\begin{itemize}
	\item Mladenovic and Hansen
	\item (seminal) A variable neighbourhood algorithm for solving the continuous location-allocation problem (1996)
	\item (overview) An introduction to Variable neighborhood search (1999)
	\item (overview?) Variable neighbourhood search: Principles and applications (2001)
	\item (overview) Developments of variable neighborhood search (2001)
	\item (overview?) Variable neighbourhood search (2002)
\end{itemize}

\begin{itemize}
	\item VARIABLE NEIGHBOURHOOD SEARCH: VNS for Training Neural Networks (chapter 5) (2006) - good overview of some of the different variations on the theme of VNS
	\item (extension) Variable Neighbourhood Search for the Global Optimization of Constrained NLPs (2005)
	\item there are many many extensions and specialization's for different problem domains: Variable neighborhood descent (VND), Reduced VNS (RVNS), Basic VNS (BVNS), Skewed VNS (SVNS), General VNS (GVNS), VN Decomposition Search (VNDS), Parallel VNS (PVNS), Primal Dual VNS (P-D VNS), Reactive VNS, Backward-Forward VNS, Best improvement VNS, Exterior point VNS, VN Simplex Search (VNSS), VN Branching, etc
	\item
\end{itemize}


% 
% Conclusions: What the reader or what thre author learned by completing this this report.
% 
\section{Conclusions}
\label{sec:conclusions}
% report
todo

% 
% Contribute
% 
\section{Contribute}
\label{sec:contribute}
% simple
Found a typo in the content or a bug in the source code? 
% advanced 
Are you an expert in this technique and know some facts that could improve the algorithm description for all?
% incentive
Do you want to get that warm feeling from contributing to an open source project? 
Do you want to see your name as an acknowledgment in print?

%  ideal
Two pillars of this effort are i) that the best domain experts are people outside of the project, and ii) that this work is wrong by default. 
% advice
Please help to make this work less wrong by emailing the author `\myreportauthor' at \url{\myreportemail} or visit the project website at \url{\myreportwebsite}.

% bibliography
\bibliographystyle{plain}
\bibliography{../bibtex}

\end{document}
% EOF