% Particle Swarm Optimization

% The Clever Algorithms Project: http://www.CleverAlgorithms.com
% (c) Copyright 2010 Jason Brownlee. Some Rights Reserved. 
% This work is licensed under a Creative Commons Attribution-Noncommercial-Share Alike 2.5 Australia License.

\documentclass[a4paper, 11pt]{article}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{url}
\usepackage[pdftex,breaklinks=true,colorlinks=true,urlcolor=blue,linkcolor=blue,citecolor=blue,]{hyperref}
\usepackage{geometry}
\usepackage[ruled, linesnumbered]{../algorithm2e}
\usepackage{listings} 
\usepackage{textcomp}
\ifx\pdfoutput\@undefined\usepackage[usenames,dvips]{color}
\else\usepackage[usenames,dvipsnames]{color}
\lstset{basicstyle=\footnotesize\ttfamily,numbers=left,numberstyle=\tiny,frame=single,columns=flexible,upquote=true,showstringspaces=false,tabsize=2,captionpos=b,breaklines=true,breakatwhitespace=true,keywordstyle=\color{blue},stringstyle=\color{ForestGreen}}
\geometry{verbose,a4paper,tmargin=25mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}

% Dear template user: fill these in
\newcommand{\myreporttitle}{Particle Swarm Optimization}
\newcommand{\myreportauthor}{Jason Brownlee}
\newcommand{\myreportemail}{jasonb@CleverAlgorithms.com}
\newcommand{\myreportwebsite}{http://www.CleverAlgorithms.com}
\newcommand{\myreportproject}{The Clever Algorithms Project\\\url{\myreportwebsite}}
\newcommand{\myreportdate}{20101115}
\newcommand{\myreportversion}{1}
\newcommand{\myreportlicense}{\copyright\ Copyright 2010 Jason Brownlee. Some Rights Reserved. This work is licensed under a Creative Commons Attribution-Noncommercial-Share Alike 2.5 Australia License.}

% leave this alone, it's templated baby!
\title{{\myreporttitle}\footnote{\myreportlicense}}
\author{\myreportauthor\\{\myreportemail}\\\small\myreportproject}
\date{\today\\{\small{Technical Report: CA-TR-{\myreportdate}-\myreportversion}}}
\begin{document}
\maketitle

% write a summary sentence for each major section
\section*{Abstract} 
% project
The Clever Algorithms project aims to describe a large number of Artificial Intelligence algorithms in a complete, consistent, and centralized manner, to improve their general accessibility. 
% template
The project makes use of a standardized algorithm description template that uses well-defined topics that motivate the collection of specific and useful information about each algorithm described.
% report
This report describes the Particle Swarm Optimization algorithm.

\begin{description}
	\item[Keywords:] {\small\texttt{Clever, Algorithms, Description, Particle, Swarm, Optimization}}
\end{description} 

\section{Introduction} 
\label{sec:intro}
% project
The Clever Algorithms project aims to describe a large number of algorithms from the fields of Computational Intelligence, Biologically Inspired Computation, and Metaheuristics in a complete, consistent and centralized manner \cite{Brownlee2010}.
% description
The project requires all algorithms to be described using a standardized template that includes a fixed number of sections, each of which is motivated by the presentation of specific information about the technique \cite{Brownlee2010a}.
% this report
This report describes the Particle Swarm Optimization algorithm.

% Name
% The algorithm name defines the canonical name used to refer to the technique, in addition to common aliases, abbreviations, and acronyms. The name is used in terms of the heading and sub-headings of an algorithm description.
\section{Name} 
\label{sec:name}
% What is the canonical name and common aliases for a technique?
% What are the common abbreviations and acronyms for a technique?
% The heading and alternate headings for the algorithm description.
Particle Swarm Optimization, PSO

% Taxonomy: Lineage and locality
% The algorithm taxonomy defines where a techniques fits into the field, both the specific subfields of Computational Intelligence and Biologically Inspired Computation as well as the broader field of Artificial Intelligence. The taxonomy also provides a context for determining the relation- ships between algorithms. The taxonomy may be described in terms of a series of relationship statements or pictorially as a venn diagram or a graph with hierarchical structure.
\section{Taxonomy}
\label{sec:taxonomy}
% To what fields of study does a technique belong?
Particle Swarm Optimization belongs to the field of Swarm Intelligence and Collective Intelligence and is a sub-field of Computational Intelligence.
% What are the closely related approaches to a technique?
Particle Swarm Optimization is related to other Swarm Intelligence algorithms such as Ant Colony Optimization and like the Genetic Algorithm it is a baseline algorithm for many variations, too numerous to list.

% Inspiration: Motivating system
% The inspiration describes the specific system or process that provoked the inception of the algorithm. The inspiring system may non-exclusively be natural, biological, physical, or social. The description of the inspiring system may include relevant domain specific theory, observation, nomenclature, and most important must include those salient attributes of the system that are somehow abstractly or conceptually manifest in the technique. The inspiration is described textually with citations and may include diagrams to highlight features and relationships within the inspiring system.
% Optional
\section{Inspiration}
\label{sec:inspiration}
% What is the system or process that motivated the development of a technique?
Particle Swarm Optimization is inspired by the social foraging behavior of some animals such as flocking behavior of birds and the schooling behavior of fish.
% Which features of the motivating system are relevant to a technique?

% Metaphor: Explanation via analogy
% The metaphor is a description of the technique in the context of the inspiring system or a different suitable system. The features of the technique are made apparent through an analogous description of the features of the inspiring system. The explanation through analogy is not expected to be literal scientific truth, rather the method is used as an allegorical communication tool. The inspiring system is not explicitly described, this is the role of the ‘inspiration’ element, which represents a loose dependency for this element. The explanation is textual and uses the nomenclature of the metaphorical system.
% Optional
\section{Metaphor}
\label{sec:metaphor}
% What is the explanation of a technique in the context of the inspiring system?
Particles in the swarm fly through an environment following the fitter members of the swarm and generally biasing their movement toward historically good areas of the environment.
% What are the functionalities inferred for a technique from the analogous inspiring system?


% Strategy: Problem solving plan
% The strategy is an abstract description of the computational model. The strategy describes the information processing actions a technique shall take in order to achieve an objective. The strategy provides a logical separation between a computational realization (procedure) and a analogous system (metaphor). A given problem solving strategy may be realized as one of a number specific algorithms or problem solving systems. The strategy description is textual using information processing and algorithmic terminology.
\section{Strategy}
\label{sec:strategy}
% What is the information processing objective of a technique?
The goal of the algorithm is to have all the particles locate the optima in a multi-dimensional hyper-volume.
% What is a techniques plan of action?
This is achieved by assigning initially random positions to all particles in the space and assign small random velocities. The algorithm is executed like a simulation, advancing the position of each particle in turn based on its velocity, the best known global position in the problem space and the best known position known to a particle. The objective function is sampled after each position update. Over time, through a combination of exploration and exploitation of known good positions in the search space, the particles cluster or converge together around an optima. 

% Procedure: Abstract computation
% The algorithmic procedure summarizes the specifics of realizing a strategy as a systemized and parameterized computation. It outlines how the algorithm is organized in terms of the data structures and representations. The procedure may be described in terms of software engineering and computer science artifacts such as pseudo code, design diagrams, and relevant mathematical equations.
\section{Procedure}
\label{sec:procedure}
% What is the computational recipe for a technique?
% What are the data structures and representations used in a technique?
Algorithm~\ref{alg:pso} provides a pseudo-code listing of the Particle Swarm Optimization algorithm for minimizing a cost function. The listing shows global interaction, although does not show local neighborhood interactions.

\begin{algorithm}[htp]
	\SetLine  

	% data
	\SetKwData{GlobalBest}{$P_{g\_best}$}
	\SetKwData{ProblemSize}{ProblemSize}
	\SetKwData{Population}{Population}
	\SetKwData{PopulationSize}{$Population_{size}$}
	\SetKwData{Particle}{$P$}
	\SetKwData{CurrentBest}{$P_{p\_best}$}
	\SetKwData{CurrentPosition}{$P_{position}$}
	\SetKwData{CurrentVelocity}{$P_{velocity}$}
	\SetKwData{CurrentCost}{$P_{cost}$}
	% functions
	\SetKwFunction{Cost}{Cost}
	\SetKwFunction{StopCondition}{StopCondition}
	\SetKwFunction{RandomPosition}{RandomPosition}
	\SetKwFunction{RandomVelocity}{RandomVelocity}	
	\SetKwFunction{UpdatePosition}{UpdatePosition}
	\SetKwFunction{UpdateVelocity}{UpdateVelocity}
	
	% I/O
	\KwIn{\ProblemSize, \PopulationSize}		
	\KwOut{\GlobalBest}
  % Algorithm

	% initialize	
	\Population $\leftarrow$ $0$\;
	\For{$i=1$ $\KwTo$ \PopulationSize} {
		\CurrentPosition $\leftarrow$ \RandomPosition{\PopulationSize}\;
		\CurrentVelocity $\leftarrow$ \RandomVelocity{}\;
		\CurrentCost $\leftarrow$ \Cost{\CurrentPosition}\;
		\CurrentBest $\leftarrow$ \CurrentPosition\;
		\If{\CurrentCost $\leq$ \GlobalBest} {
			\GlobalBest $\leftarrow$ \CurrentBest\;
		}
	}	

	% loop
	\While{$\neg$\StopCondition{}} {
		\ForEach{\Particle $\in$ \Population} {
			\CurrentVelocity $\leftarrow$ \UpdateVelocity{\CurrentVelocity, \GlobalBest, \CurrentBest}\;
			\CurrentPosition $\leftarrow$ \UpdatePosition{\CurrentPosition, \CurrentVelocity}\;
			\CurrentCost $\leftarrow$ \Cost{\CurrentPosition}\;
			\If{\CurrentCost $\leq$ \CurrentBest} {
				\CurrentBest $\leftarrow$ \CurrentPosition\;
				\If{\CurrentCost $\leq$ \GlobalBest} {
					\GlobalBest $\leftarrow$ \CurrentBest\;
				}
			}
		}		
	}
	\Return{\GlobalBest}\;
	% end
	\caption{Pseudo Code for the Particle Swarm Optimization algorithm (global max).}
	\label{alg:pso}
\end{algorithm}


% Heuristics: Usage guidelines
% The heuristics element describe the commonsense, best practice, and demonstrated rules for applying and configuring a parameterized algorithm. The heuristics relate to the technical details of the techniques procedure and data structures for general classes of application (neither specific implementations not specific problem instances). The heuristics are described textually, such as a series of guidelines in a bullet-point structure.
\section{Heuristics}
\label{sec:heuristics}
% What are the suggested configurations for a technique?
% What are the guidelines for the application of a technique to a problem instance?
\begin{itemize}
	\item The number of particles should be low, around 20-40
	\item The speed a particle can move (maximum change is position per iteration) should be bounded, such as to a percentage of the size of the domain.
	\item The learning factors (biases towards global and personal best positions) should be between 0 and 4, typically 2.
	\item A local bias (local neighborhood) factor can be introduced where neighbors are determined based on Euclidean distance between particle positions.
	\item Particles may leave the boundary of the problem space and may be penalized or be reflected back into the domain. 
\end{itemize}

% The code description provides a minimal but functional version of the technique implemented with a programming language. The code description must be able to be typed into an appropriate computer, compiled or interpreted as need be, and provide a working execution of the technique. The technique implementation also includes a minimal problem instance to which it is applied, and both the problem and algorithm implementations are complete enough to demonstrate the techniques procedure. The description is presented as a programming source code listing.
\section{Code Listing}
\label{sec:code}
% How is a technique implemented as an executable program?
% How is a technique applied to a concrete problem instance?
Listing~\ref{pso} provides an example of the Particle Swarm Optimization algorithm implemented in the Ruby Programming Language. 
% problem
The demonstration problem is an instance of a continuous function optimization that seeks $min f(x)$ where $f=\sum_{i=1}^n x_{i}^2$, $-5.0\leq x_i \leq 5.0$ and $n=3$. The optimal solution for this basin function is $(v_0,\ldots,v_{n-1})=0.0$.
% algorithm
TODO

% the listing
\lstinputlisting[firstline=7,language=ruby,caption=Particle Swarm Optimization in the Ruby Programming Language, label=pso]{../../src/algorithms/swarm/pso.rb}


% References: Deeper understanding
% The references element description includes a listing of both primary sources of information about the technique as well as useful introductory sources for novices to gain a deeper understanding of the theory and application of the technique. The description consists of hand-selected reference material including books, peer reviewed conference papers, journal articles, and potentially websites. A bullet-pointed structure is suggested.
\section{References}
\label{sec:references}
% What are the primary sources for a technique?
% What are the suggested reference sources for learning more about a technique?

% 
% Primary Sources
% 
\subsection{Primary Sources}
% seminal
Particle Swarm Optimization was described as a stochastic global optimization method for continuous functions in 1995 by Eberhart and Kennedy \cite{Eberhart1995, Kennedy1995}. It was based on work on genetic algorithms and on the flocking behavioral models of Reynolds \cite{Reynolds1987}.
% early
Early works included the introduction of inertia \cite{Shi1998} and early study of social topologies in the swarm by Kennedy \cite{Kennedy1999}. 

% 
% Learn More
% 
\subsection{Learn More}
% reviews

Poli provides a meta-analysis of PSO publications that focus on the application the technique, providing a systematic breakdown on application areas \cite{Poli2008a}. 
% books
An excellent book on Swarm Intelligence in general with detailed coverage of Particle Swarm Optimization is ``Swarm Intelligence'' by Kennedy, Eberhart, and Shi \cite{Kennedy2001}.

% 
% Conclusions: What the reader or what thre author learned by completing this this report.
% 
\section{Conclusions}
\label{sec:conclusions}
% report
todo

% 
% Contribute
% 
\section{Contribute}
\label{sec:contribute}
% simple
Found a typo in the content or a bug in the source code? 
% advanced 
Are you an expert in this technique and know some facts that could improve the algorithm description for all?
% incentive
Do you want to get that warm feeling from contributing to an open source project? 
Do you want to see your name as an acknowledgment in print?

%  ideal
Two pillars of this effort are i) that the best domain experts are people outside of the project, and ii) that this work is subjected to continuous improvement. 
% advice
Please help to make this work less wrong by emailing the author `\myreportauthor' at \url{\myreportemail} or visit the project website at \url{\myreportwebsite}.

% bibliography
\bibliographystyle{plain}
\bibliography{../bibtex}

\end{document}
% EOF