% Technical Report: Algorithm Testing

% The Clever Algorithms Project: http://www.CleverAlgorithms.com
% (c) Copyright 2010 Jason Brownlee. Some Rights Reserved. 
% This work is licensed under a Creative Commons Attribution-Noncommercial-Share Alike 2.5 Australia License.

\documentclass[a4paper, 11pt]{article}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{url}
\usepackage[pdftex,breaklinks=true,colorlinks=true,urlcolor=blue,linkcolor=blue,citecolor=blue,]{hyperref}
\usepackage{geometry}
\usepackage[ruled, linesnumbered]{../algorithm2e}
\usepackage{listings} 
\usepackage{textcomp}
\ifx\pdfoutput\@undefined\usepackage[usenames,dvips]{color}
\else\usepackage[usenames,dvipsnames]{color}
\lstset{basicstyle=\footnotesize\ttfamily,numbers=left,numberstyle=\tiny,frame=single,columns=flexible,upquote=true,showstringspaces=false,tabsize=2,captionpos=b,breaklines=true,breakatwhitespace=true,keywordstyle=\color{blue},stringstyle=\color{ForestGreen}}
\geometry{verbose,a4paper,tmargin=25mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}

% Dear template user: fill these in
\newcommand{\myreporttitle}{Clever Algorithms}
\newcommand{\myreportsubtitle}{Unit Tests and Writing for Testability}
\newcommand{\myreportauthor}{Jason Brownlee}
\newcommand{\myreportemail}{jasonb@CleverAlgorithms.com}
\newcommand{\myreportproject}{The Clever Algorithms Project\\\url{http://www.CleverAlgorithms.com}}
\newcommand{\myreportdate}{20101220}
\newcommand{\myreportfulldate}{December 20, 2010}
\newcommand{\myreportversion}{1}
\newcommand{\myreportlicense}{\copyright\ Copyright 2010 Jason Brownlee. Some Rights Reserved. This work is licensed under a Creative Commons Attribution-Noncommercial-Share Alike 2.5 Australia License.}

% leave this alone, it's templated baby!
\title{{\myreporttitle}: {\myreportsubtitle}\footnote{\myreportlicense}}
\author{\myreportauthor\\{\myreportemail}\\\small\myreportproject}
\date{\myreportfulldate\\{\small{Technical Report: CA-TR-{\myreportdate}-\myreportversion}}}
\begin{document}
\maketitle

% write a summary sentence for each major section
\section*{Abstract} 
% project
The Clever Algorithms project aims to describe a large number of Artificial Intelligence algorithms in a complete, consistent, and centralized manner, to improve their general accessibility. 
% template
The project makes use of a standardized algorithm description template that uses well-defined topics that motivate the collection of specific and useful information about each algorithm described.
% report
This report provides an introduction to software testing and the testing of Artificial Intelligence algorithms

\begin{description}
	\item[Keywords:] {\small\texttt{Clever, Algorithms, Test, Unit Test, Testability}}
\end{description} 

% summarise the document breakdown with cross references
\section{Introduction}
\label{sec:introduction}
% project
The Clever Algorithms project aims to describe a large number of algorithms from the fields of Computational Intelligence, Biologically Inspired Computation, and Metaheuristics in a complete, consistent and centralized manner \cite{Brownlee2010}.
% description
The project requires all algorithms to be described using a standardized template that includes a fixed number of sections, each of which is motivated by the presentation of specific information about the technique \cite{Brownlee2010a}.

% report
This report provides an introduction to software testing and the testing of Artificial Intelligence algorithms. Section~\ref{sec:testing} introduces software testing and focuses on a type of testing relevant to testing algorithms called unit testing. Section~\ref{sec:example} provides a specific example of an algorithm and a prepared suite of unit tests, and Section~\ref{sec:suggestions} provides some rules-of-thumb for testing algorithms in general.


%
% Software Testing
%
\section{Software Testing}
\label{sec:testing}
% software
Testing is the field of Software Engineering is a process in the life-cycle of a software product or service that verifies that the software meets quality expectations validates that software meets the requirements specification. Software testing is intended to locate defects in a program, although a given testing method cannot grantee to locate all defects. As such, it is common for an application to be subjected to a range of testing methodologies throughout the software life-cycle, such as unit testing during development, integration system testing once modules and systems are completed, and user acceptance testing to allow the stakeholders to be satisfied that needs are met.

% unit testing
Unit testing is a type of software testing that involves the preparation of well-defined procedural tests of discrete functionality of a program that provide confidence that a module or function behaves as intended. Unit tests are referred to as `white-box' tests (contrasted to `black-box' tests) given that they are written with full knowledge of the internal structure of the functions and modules under tests. Unit tests are typically prepared by the developer that wrote the code under test and are commonly automated, themselves written as small programmers that are executed by a unit testing framework (such as JUnit for Java). The objective is not to test each path within a function (called complete-test or complete-code coverage), but instead to focus tests on areas of risk, uncertainty, or criticality. Each test should focus on one aspect of the code (test one thing) and are organized into test suites of commonality. 

Some benefits of unit testing include:

\begin{itemize}
  \item \emph{Documentation} The preparation of a suite of tests for a given system provide a type of programming documentation highlighting the expected behavior of functions and modules and providing examples of how to interact with key components.
  \item \emph{Readability} Unit testing encourages a programming style of small modules, clear input and output and possibly fewer inter-component dependencies. Code written for easy of testing (testability) may be easier to read and follow by third parties.
  \item \emph{Regression} Together, the suite of tests can be executed as a regression-test of the system in the future after changes to the code have been made. The automation of the tests means that any defects caused by changes to the code can easily be identified.
\end{itemize}

% test driven development
Unit tests were traditionally written after the program was completed. An modern alternative is to prepare the tests before the functionality of the application is prepared, called Test-First or Test-Driven Development (TDD). In this method, the tests are written and execute, failing until the application functionality is written to make the test pass. The early preparation of tests allow the programmer to consider the behavior required from the program and the interfaces or functions with which the program needs to expose before they are written. 

% algorithm testing
The concerns of software testing are very relevant to the development, investigation, and application of Artificial Intelligence algorithms. In particular the strong culture of empirical investigation and prototype-based development demand a baseline level of trust in the systems that are presented in articles and papers. Trust can be instilled in an algorithm by assessing the quality of the algorithm implementation itself. 
% testing algorithms
Unit testing is light weight (requiring only the writing of automated test code) and meets the needs of promoting quality and trust in prototyping and developing algorithms. It is strongly suggested as a necessary step in the process of empirical algorithm research in the fields of Metaheuristics, Computational Intelligence, and Biologically Inspired Computation.

%
% Example
%
\section{Unit Testing Example}
\label{sec:example}
This section provides an example of an algorithm and its associated unit tests as an illustration of the presented concepts. Section~\ref{subsec:algorithm} presents the Genetic Algorithm and Section~\ref{subsec:tests} presents the unit tests for the Genetic Algorithm, both written in the Ruby Programming Language.

% Algorithm
\subsection{Algorithm}
\label{subsec:algorithm}
Listing~\ref{ga} presents the source code for the Genetic Algorithm in the Ruby Programming language. An important consideration in using the Ruby test suite, is ensuring that functions of the algorithm are exposed for testing and that the algorithm example itself does not execute. This is achieved through the use of the \texttt{if \_\_FILE\_\_ == \$0} condition that ensure the example only executes when called directly, allowing the functions to be imported and executed independently by a unit test script.

The Genetic Algorithm is used as the basis for example as it is well understood ad provides an implementation that is an excellent case for demonstrating unit testing methodology \cite{Brownlee2010p}. The algorithm is very modular with its behavior partitioned into well-contained functions, most which are independently testable. 

The \texttt{reproduce} has some dependencies although is still testable. The \texttt{search} function is the only monolithic function, which both depends on all other functions in the implementation (directly or indirectly) and is difficult to unit test. At best the \texttt{search} function may be a case for system testing addressing functional requirements, such as ``\emph{does the algorithm deliver optimized solutions}''.

% the listing
\lstinputlisting[firstline=7,language=ruby,caption=Genetic Algorithm in the Ruby Programming Language, label=ga]{../../src/algorithms/evolutionary/genetic_algorithm.rb}

% Unit Test
\subsection{Unit Tests}
\label{subsec:tests}
Listing~\ref{gatests} provides the \texttt{TC\_GeneticAlgorithm} class that makes use of the built-in Ruby unit testing framework by extending the \texttt{Test::Unit::TestCase} class.
The listing provides an example of 10 unit tests for 6 of the 7 functions in the Genetic Algorithm implementation. Two types of unit tests are provided: 

\begin{itemize}
	\item \emph{Deterministic}: Tests that directly test the function in question, addressing questions such as: does \texttt{onemax} add correctly? and does \texttt{point\_mutation} behave correctly?
	\item \emph{Probabilistic}: Tests that test the probabilistic properties of the function in question, addressing questions such as: does \texttt{random\_bitstring} provide an expected 50/50 mixture of $1$'s and $0$'s over a large number of cases? and does \texttt{point\_mutation} make an expected number of changes over a large number of cases?
\end{itemize}

This second type that tests for probabilistic expectations is a weaker form of unit testing that can be used to either provide additional confidence to directly tested functions, or to be used as a last resort where direct methods cannot be used.

Given that a unit test should `test one thing' it is common for a given function to have more than one unit tests. The \texttt{reproduce} function is a good example of this with three tests in the suite. This is because it is a larger function with behavior called in dependent functions and is varied based on parameters.

% the listing
\lstinputlisting[firstline=10,language=ruby,caption=Unit Tests for the Genetic Algorithm the Ruby Programming Language, label=gatests]{../../src/algorithms/evolutionary/tests/tc_genetic_algorithm.rb}

%
% Rules-of-Thumb
%
\section{Rules-of-Thumb}
\label{sec:suggestions}
Unit testing is easy, although writing good unit tests is difficult given the complex relationship the tests have with the code under test. Testing Metaheuristics and Computational Intelligence algorithms is harder again given their probabilist nature and their ability to `work in-spite of you', that is, provide some kind of result even when implemented with defects.

The following provide 10 guidelines that may help when it comes time to unit testing an algorithm:

\begin{itemize}
 	\item \emph{Start Small}: Some unit tests is better than no unit test and each additional test can improve the trust and potentially the quality of the code. For an existing algorithm implementation, start by writing a test for a small and simple behavior and slowly build up a test suite.
  \item \emph{Test one thing}: Each test should focus on verifying the behavior of one aspect of one unit of code. Concise unit tests are brief unit tests.
  \item \emph{Test once}: A behavior or expectation only needs to be tested once, do not repeat a test each time a given unit is tested.
  \item \emph{Don't forget the I//O}: Remember to test the inputs and outputs of a unit of code, specifically the pre-conditions and post-conditions. It can be easy to focus on the decision points within a function and forget its primary purpose.
	\item \emph{Write code for testability}: The tests should help to shape the code they test. Write small functions or modules, think about testing while writing code (or write tests first), and refactor code (update code after the fact) to make it easier to test.
	\item \emph{Function independence}: Attempt to limit the direct dependence between functions, modules, objects and other constricts. This is related to testability and writing small functions although suggests limits on how much interaction there is between units of code in the algorithm. Less dependence means less side-effects of a given unit of code and ultimately complicated tests.
  \item \emph{Test Independence}: Test should be independent from each other. Frameworks provide hooks to set-up and tear-down state prior to the execution of each test, there should be no needed to have one test prepare data or state for another tests. Tests should be able to execute independently and in any order.
	\item \emph{Test your own code}: Avoid writing tests that verify the behavior of framework or library code such as the randomness of a random number generator or whether a math or string function behaves as expected. Focus on writing test for the manipulation of data performed by the code you have written.
	\item \emph{Probabilistic testing}: Metaheuristics and Computational Intelligence algorithms generally make use of stochastic or probabilistic decisions. This means that some behaviors are not deterministic and are more difficult to test. As with the example, write probabilistic tests to verify that such processes behave as intended.
	\item \emph{Consider test-first}: Writing the tests first in a Test-Driven methodology can help to crystallize expectations when implementing an algorithm from the literature, and help to solidify thoughts when developing or prototyping a new idea.
\end{itemize}


% summarise the document message and areas for future consideration
\section{Conclusions}
\label{sec:conclusions}
This report provided an introduction to software testing, unit testing and provided heuristics for unit testing Metaheuristic and Computational Intelligence algorithms with examples.


% bibliography
\bibliographystyle{plain}
\bibliography{../bibtex}

\end{document}
% EOF
