\chapter{Stochastic Algorithms}
\label{ch:stochastic}

This chapter is all about stochastic algorithms.

\section{Overview}

% chapter introduction
An important base principle of computational intelligence algorithm is non-determinism. That is, algorithms that make stochastic and probabilistic decisions and ultimately produce varied outputs based on the same inputs. This chapter introduces some basic stochastic tools that are exploited by most if not all inspired algorithms in this book. The tools are demonstrated with some stochastic-based search algorithms.

\subsection{Global Optimization}
randomization methods, classical stuff

% consider moving the random number generation stuff into this section

\subsection{Further Reading}
Where to read more about random numbers and search / optimizations

% sections
\input{stochastic/numbers}
\input{stochastic/randomsearch}
\input{stochastic/localizedrandomsearch}
\input{stochastic/iteratedlocalsearch}
\input{stochastic/adaptiverandomsearch}
\input{stochastic/multiplerestart}
\input{stochastic/tabusearch}
\input{stochastic/grasp}
\input{stochastic/stochasticdiffusion}

\section{Summary}
This chapter introduced a number of important principles that underlie all inspired algorithms, as follows:

\begin{itemize}
	\item Pseudorandom numbers as a source of non-determinism.
	\item Randomization as a method of exploration and discovery.
	\item Biasing a sampling procedure to increase efficiency and efficacy.
	\item Multiple restarts to address convergence.	
\end{itemize}

