% 
% Errata
% 

\chapter*{Errata\markboth{Errata}{}}
\addcontentsline{toc}{part}{Errata}

\section*{Revision 2}
\begin{small}
	
	\begin{description}
		\item[page 43--47] Typo's of Multi-Restart. Thanks to Stephan Williams.
		\item[page 307] Confusion of the role of dendrites and axons for a neuron. Thanks to Ricardo Colome.
		\item[page 311] Confusion of the role of dendrites and axons for a neuron. Thanks to Ricardo Colome.
		\item[page 316] Confusion of the role of dendrites and axons for a neuron. Thanks to Ricardo Colome.
	\end{description}

\end{small}


\section*{Revision 1}

\begin{small}
	
\begin{description}
	\item[page 9] Typo in Metaheuristics section of the Introduction. Thanks to Leif Wickland.
	\item[page 11] Typo in Function Optimization section of the Introduction. Thanks to John Wise and Patrick Boehnke.
	\item[page 11] Typo's in the Function Approximation section of the Introduction. Thanks to Patrick Boehnke.
	\item[page 13] Typo in the Function Approximation section of the Introduction. Thanks to Patrick Boehnke.
	\item[page 32] Typo in References section of Random Search.
  \item[page 37] Fixed bug with \texttt{step\_size} in Adaptive Random Search implementation. Thanks to Zach Scott.
	\item[page 43] Typo in Taxonomy section of Iterated Local Search. Thanks to Diego Noble.
	\item[page 69] Bug in \texttt{recombine} function of the Scatter Search algorithm. Thanks to Markus Stokmaier.
	\item[page 111] Bug in the \texttt{init\_population} function of the Evolution Strategies algorithm. Thanks to Lai Yu-Hsuan.
	\item[page 129] Bug in the \texttt{one\_point\_crossover} function of the Grammatical Evolution implementation. Thanks to Mark Chenoweth.
	\item[page 234] Fixed ambiguous pseudo code description of Particle Swarm Optimization. Thanks to Stefan Pauleweit.
	\item[page 235] Fixed a bug in the \texttt{get\_global\_best} function of the Particle Swarm Optimization implementation. Thanks to Paul Chinnery.
	\item[page 237] Improved reference 3 for Particle Swarm Optimization. Thanks to Diego Noble.
	\item[page 242] Fixed a bug in the \texttt{search} function of the Ant System implementation. Thanks to Andrew Myers.
	\item[page 330] Typo in taxonomy of LVQ algorithm. Thanks to Jason Davies.
	\item[page 393] Typo in Function Approximation section. Thanks to Diego Noble.
	\item[page 400] Typo in subsection 9.6.1. Thanks to Diego Noble.
	\item[page 402] Typo in subsection 9.6.2. Thanks to Diego Noble.
	\item[page 415] Changed equality to assignment in Ruby flow control example in Appendix A. Thanks to Donald Doherty.
	\item[page 413] Typo in Overview section in Appendix A. Thanks to Martin-Louis Bright.
	\item[page 413] Typo in Ruby Files section in Appendix A. Thanks to Brook Tamir.
	\item[page 414] Typos in Variables section in Appendix A. Thanks to Brook Tamir.
	\item[page 415] Typos in Flow Control and Arrays sections in Appendix A. Thanks to Brook Tamir.
	\item[page 416] Typos in Arrays and Function and Blocks sections in Appendix A. Thanks to Brook Tamir.
	\item[page 417] Typos in Function and Blocks section in Appendix A. Thanks to Brook Tamir.
	\item[page 418] Typos in Enumerating section in Appendix A. Thanks to Brook Tamir.
	\item[page 419] Typo in Conclusions section in Appendix A. Thanks to Brook Tamir.
\end{description}

\end{small}
